---
title: "STAT430_HW01"
author: "Taiga Hasegawa(taigah2)"
date: "2019/1/25"
output: pdf_document
---
---
title: "STAT430"
author: "Taiga Hasegawa(taigah2)"
date: "2019/1/18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
amazon = read.csv("AMZN_2012-06-21_34200000_57600000_message_10.csv", header = F)
names(amazon)=c("Time" , "Type", "OrderID", "Size", "Price", "Direction")
amazon$Size <- as.numeric(amazon$Size)
amazon$Price <- as.numeric(amazon$Price)
```

```{r}
library("lubridate")
```
```{r}
demodate="2016-06-21"
options(digits.secs=3)
amazon$tstamp=as_datetime(demodate,tz="US/Eastern")+amazon$Time
```

```{r}
amazon_subset=subset(amazon, Type %in% c(4,5))
head(amazon_subset,n=5)
```

```{r}
imbalance_tick <- function(dat)
{
  n <- length(dat$Price)
  imbalance <- rep(0, n)
  price_diff <- diff(dat$Price)
  for(i in 2:n)
  {
    imbalance[i] <- sign(price_diff[i-1])*(price_diff[i-1]!=0) + imbalance[i-1]*(price_diff[i-1]==0)
  }
  imbalance
}
```



```{r}
library(pracma)
Tstar_tib=function(data,w0=10,bkw_T=5,bkw_b=5){
  b_t=imbalance_tick(data)
  w0=max(min(which(cumsum(b_t)!=0)),w0)
  Tstar=w0
  E0t=Tstar
  repeat{
    Tlast=sum(Tstar)
    nbt=min(bkw_T,Tlast-1)
    P=pracma::movavg(b_t[1:Tlast],n=nbt,type="e")
    P<- tail(P,1) 
    b_t_expected=E0t*abs(P)
    b_t_cumsum=abs(cumsum(b_t[-(1:Tlast)]))
    if(max(b_t_cumsum)<b_t_expected){break}else{
      Tnew=min(which(b_t_cumsum>=b_t_expected))
    }
    Tlast=Tlast+Tnew
    if(Tlast>dim(data)){break}else{
      Tstar=c(Tstar,Tnew)
      if(length(Tstar)<=2){
        E0t=mean(Tstar)
      }else{
        nt=min(bkw_T,length(Tstar)-1)
        E0t=pracma::movavg(Tstar[1:length(Tstar)],n=nt,type="e")
        E0t <- tail(E0t,1)
      }
    }
  }
  return(Tstar)
}
```

```{r}
##Tick imbalance bar
tib=Tstar_tib(amazon_subset)
```


```{r}
imbalance_volume <- function(dat)
{
  n <- length(dat$Price)
  imbalance <- rep(0, n)
  b_t=rep(0,n)
  price_diff <- diff(dat$Price)
  v_t=dat$Size
  for(i in 2:n)
  {
    imbalance[i]=sign(price_diff[i-1])*(price_diff[i-1]!=0)*v_t[i]+imbalance[i-1]*(price_diff[i-1]==0)
  }
  imbalance
}
```

```{r}
Vstar_vib=function(data,w0=10,bkw_T=5,bkw_b=5){
  bv_t=imbalance_volume(data)
  w0=max(min(which(cumsum(bv_t)!=0)),w0)
  Tstar=w0
  E0t=Tstar
  repeat{
    Tlast=sum(Tstar)
    nbt=min(bkw_b,Tlast-1)
    P=pracma::movavg(bv_t[1:Tlast],n=nbt,type="e")
    P=tail(P,1)
    bv_t_expected=E0t*abs(P)
    bv_t_cumsum=abs(cumsum(bv_t[-(1:Tlast)]))
    if(max(bv_t_cumsum)<bv_t_expected){break}else{
      Tnew=min(which(bv_t_cumsum>=bv_t_expected))
    }
    Tlast=Tlast+Tnew
    if(Tlast>dim(data)[1]){break}else{
      Tstar=c(Tstar,Tnew)
      if(length(Tstar)<=2){
        E0t=mean(Tstar)
      }else{
        nt=min(bkw_T,length(Tstar)-1)
        E0t=pracma::movavg(Tstar[1:length(Tstar)],n=nt,type="e")
        E0t=tail(E0t,1)
      }
    }
  }
  return(Tstar)
}
```

```{r}
##Volume imbalance bar
vib=Vstar_vib(amazon_subset)
cumsum(vib)
```

```{r}
Tstar_trb <- function(dat, w0=10, bkw_T=5, bkw_Pb1=5)
{
  b_t <- imbalance_tick(dat)
  nb <- length(b_t)
  nx <- dim(dat)[1]
  th_T <- sapply(1:nb, function(i){
    b_t_tmp <- b_t[1:i]
    if(sum(b_t_tmp %in% c(-1,1))==0){out <- 0}else
    {
      out <- max(cumsum(b_t_tmp[b_t_tmp==1]), -cumsum(b_t_tmp[b_t_tmp==-1]))  
    }
    out
  })
  w0 <- max(min(which(th_T != 0)), w0) 
  w0 <- max(min(which(b_t==1)), w0) 
  Tvec <- w0
  E0T <- T_last <- Tvec
  Pb1 <- sum(b_t[1:w0]==1) / w0 
  Pb1vec <- Pb1
  th_T_Expected <- E0T*max(Pb1, 1-Pb1)
  while(T_last<nx)
  {
    T_last <- sum(Tvec) 
    for(j in 1:(nb-T_last-1))
    {
      b_t_tmp <- b_t[(T_last+1):(T_last+j)]
      if(sum(b_t_tmp %in% c(-1,1))==0){th_T_tmp <- 0}else
      {
        th_T_tmp <- max(cumsum(b_t_tmp[b_t_tmp==1]), -cumsum(b_t_tmp[b_t_tmp==-1]))
      }
      if(th_T_tmp >= th_T_Expected)
      {
        new_flag <- TRUE 
        T_new <- j
        Tvec <- c(Tvec, T_new)
        T_last <- T_last + T_new
        
        Pb1_new <- sum(b_t_tmp==1) / j
        Pb1vec <- c(Pb1vec, Pb1_new)
        break
      }
    }
    
    if(new_flag==TRUE)
    {
      new_flag <- FALSE
      nTvec <- length(Tvec) 
      if(nTvec <= 2)
      {
        E0T <- mean(Tvec) 
        Pb1 <- mean(Pb1vec)
      }else
      {
        nT <- min(bkw_T, length(Tvec)-1) 
        E0T <- pracma::movavg(Tvec[(nTvec-nT):nTvec], n=nT, type = "e")
        E0T <- E0T[length(E0T)]
        nPb1 <- min(bkw_Pb1, length(Tvec)-1)
        Pb1 <- pracma::movavg(Pb1vec[(nTvec-nPb1):nTvec], n=nPb1, type = "e")
        Pb1 <- Pb1[length(Pb1)]
      }
      th_T_Expected <- E0T*max(Pb1, 1-Pb1)
    }else{break}
  }
  return(Tvec)
}
```

```{r}
##Tick runs bar
trb=Tstar_trb(amazon_subset)
```


```{r}
Vstar_vrb <- function(dat, w0=10, bkw_T=5, bkw_Pb1=5, bkw_v=5)
{
  b_t <- imbalance_tick(dat)
  nb <- length(b_t)
  nx <- dim(dat)[1]
  volume=amazon_subset$Size
  th_T <- sapply(1:nb, function(i){
    b_t_tmp <- b_t[1:i]
    volume_tmp=volume[1:i]
    if(sum(b_t_tmp %in% c(-1,1))==0){out <- 0}else
    {
      index1=which(b_t_tmp==1)
      index2=which(b_t_tmp==-1)
      index_for_1=c(index1)
      index_for_2=c(index2)
      out <- max(cumsum(b_t_tmp[index1]*volume_tmp[index1]), -cumsum(b_t_tmp[index2]*volume_tmp[index2]))
    }
    out
  })
  
  w0 <- max(min(which(th_T != 0)), w0) 
  w0 <- max(min(which(b_t==1)), w0) 
  Tvec <- w0
  E0T <- T_last <- Tvec
  Pb1 <- sum(b_t[1:w0]==1) / w0 
  Pb1vec <- Pb1
  bool1=b_t[1:w0]==1
  nw1=min(bkw_v,sum(bool1))
  volume_average_1=pracma::movavg(volume[bool1],n=nw1,type = "e")
  volume_average_1=volume_average_1[length(volume_average_1)]
  bool2=b_t[1:w0]==-1
  nw2=min(bkw_v,sum(bool2))
  volume_average_2=pracma::movavg(volume[bool2],n=nw2,type = "e")
  th_T_Expected <- E0T*max(Pb1*volume_average_1, (1-Pb1)*volume_average_2)
  volume_average_2=volume_average_2[length(volume_average_2)]
  
  while(T_last<nx)
  {
    T_last <- sum(Tvec) 
    for(j in 1:(nb-T_last-1))
    {
      b_t_tmp <- b_t[(T_last+1):(T_last+j)]
      volume_tmp=volume[(T_last+1):(T_last+j)]
      if(sum(b_t_tmp %in% c(-1,1))==0){th_T_tmp <- 0}else
      {
        index1=which(b_t_tmp==1)
        index2=which(b_t_tmp==-1)
        th_T_tmp <- max(cumsum(b_t_tmp[index1]*volume_tmp[index1]), -cumsum(b_t_tmp[index2]*volume_tmp[index2]))
      }
      if(th_T_tmp >= th_T_Expected)
      {
        new_flag <- TRUE 
        T_new <- j
        Tvec <- c(Tvec, T_new)
        T_last <- T_last + T_new
        
        Pb1_new <- sum(b_t_tmp==1) / j
        Pb1vec <- c(Pb1vec, Pb1_new)
        break
      }
    }
    
    if(new_flag==TRUE)
    {
      new_flag <- FALSE
      nTvec <- length(Tvec) 
      bool1=b_t[1:T_last]==1
      nw1=min(bkw_v,sum(bool1))
      volume_average_1=pracma::movavg(volume[bool1],n=nw1,type = "e")
      volume_average_1=volume_average_1[length(volume_average_1)]
      bool2=b_t[1:T_last]==-1
      nw2=min(bkw_v,sum(bool2))
      volume_average_2=pracma::movavg(volume[bool2],n=nw2,type = "e")
      volume_average_2=volume_average_2[length(volume_average_2)]
      if(nTvec <= 2)
      {
        E0T <- mean(Tvec)
        Pb1 <- mean(Pb1vec)
      }else
      {
        nT <- min(bkw_T, length(Tvec)-1) 
        E0T <- pracma::movavg(Tvec[(nTvec-nT):nTvec], n=nT, type = "e")
        E0T <- E0T[length(E0T)]
        nPb1 <- min(bkw_Pb1, length(Tvec)-1)
        Pb1 <- pracma::movavg(Pb1vec[(nTvec-nPb1):nTvec], n=nPb1, type = "e")
        Pb1 <- Pb1[length(Pb1)]
      }
      th_T_Expected <- E0T*max(Pb1*volume_average_1, (1-Pb1)*volume_average_2)
    }else{break}
  }
  return(Tvec)
}
```

```{r}
vrb=Vstar_vrb(amazon_subset)
```


```{r}
plot(amazon_subset$Price, pch=20, xlab="ticks", ylab="Price", main="Where to sample tick imbalance bars?")
abline(v=cumsum(tib), lwd=0.2)
```

```{r}
plot(amazon_subset$Price, pch=20, xlab="ticks", ylab="Price", main="Where to sample volume imbalance bars?")
abline(v=cumsum(vib), lwd=0.2)
```

```{r}
plot(amazon_subset$Price, pch=20, xlab="ticks", ylab="Price", main="Where to sample tick runs bars?")
abline(v=cumsum(trb), lwd=0.2)
```

```{r}
plot(amazon_subset$Price, pch=20, xlab="ticks", ylab="Price", main="Where to sample volume runs bars?")
abline(v=cumsum(vrb), lwd=0.2)
```
Tick imbalance bars show up at 10, 28, 42, 105, 284, 818 and 2680 ticks. Volume imbalance bars show up at 10, 42, 55, 245, and 724ticks. They have the some same values but volume imbalance bars have fewer values. 
Tick rus bars and volume runs bars have much more bars than imbalance bars. The number of bars are increasing when the price goes up and down in both runs bars. Tick runs bars have more bars than volume runs bars. 
They are all made for detecting the unusual deviation from the normal price or volume. Runs bars are more useful than imbalance bars because the run bars shows the bars more obviously when the price goes up and down.  
